---
title: "Homework 1"
author: "Lina Marcinczyk"
date: "2026-09-21"
output: github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
library(tidyverse) 
```

## Problem 1


```{r}
data("penguins", package = "palmerpenguins")
```

### Description of the pengiuns dataset

The `penguins` dataset contains measurements on penguins from three species:
`r paste(unique(penguins$species), collapse = ", ")`.
Important variables include species, island, bill length, bill depth,
flipper length, body mass, sex, and year.

The dataset contains `r nrow(penguins)` rows and `r ncol(penguins)` columns

The mean flipper length is `r mean(penguins$flipper_length_mm, na.rm = TRUE)` mm.

### Scatterplot

```{r}

penguin_plot =
  ggplot(
    penguins,
    aes(
      x = bill_length_mm,
      y = flipper_length_mm,
      color = species
    )
  ) +
  geom_point()

penguin_plot
```

## Problem 2

```{r}
problem2_df =
  tibble(
    random_sample = rnorm(10),
    sample_positive = random_sample > 0,
    character_vector = c(
      "a", "b", "c", "d", "e",
      "f", "g", "h", "i", "j"
    ),
    factor_vector = factor(
      c(
        "A", "B", "C", "A", "B",
        "C", "A", "B", "C", "A"
      )
    )
  )

problem2_df
```

```{r}
mean(pull(problem2_df, random_sample))
mean(pull(problem2_df, sample_positive))
mean(pull(problem2_df, character_vector))
mean(pull(problem2_df, factor_vector))
```


